My work has been possible with the mentorship and support of many people and I now take the opportunity to thank them and convey my heartfelt gratitude.

I thank my thesis advisor Prof. Stephen Parker for his exceptional mentorship. He's a brilliant scientist, creative and ambitious and is a genuinely nice person. It has been great working with and learning from him. He looks out for his students, gives constructive feedback and is one of the most optimistic people I know. Of all the things I have learned from him, staying positive is something I'll especially continue to strive for in all walks of life. I thank my fellow lab members Ricardo Albanus, Dr. Venkat Elangovan, John Hensley, Dr. Yoshi Kyono, Dr. Nandini Manickam, Peter Orchard, Dr. Daniel Quang and Vivek Rai for their support, motivation and friendship. It has been an honor to have known them and worked with them. 

My thesis committee comprised by Prof. Sally Camper, Prof. Jacob Kitzman, Prof. Mats Ljungman and Prof. Cristen Willer has been just wonderful. All members have given invaluable feedback and support throughout my dissertation work. I thank them for helping me rethink goals and approaches for my final project and helping me stay motivated. I always felt better about my work after my committee meetings. I also thank Jacob for his contributions towards experiments and analyses for my project.

I have been extremely lucky to have had opportunities to work with so many brilliant and kind people through the FUSION, insPIRE and MAGIC groups. Not only did I learn scientific skills from them, but I also acquired other critical skills such as teamwork and communication. I especially thank Dr. Mike Erdos, Narisu Narisu, Dr. Laura Scott, Dr. Ana Vi\~{n}uela, Dr. Martijn van de Bunt, Prof. In\^{e}s Barroso, Prof. Mark McCarthy, Prof. Karen Mohlke, Prof. Mike Boehnke and Prof. Francis Collins for their kindness and support. It has been my absolute honor to have worked with them all. I especially thank Prof. Karen Mohlke and Prof. Francis Collins for helping with letters of recommendation. 
                                
I thank the Human Genetics (HG) department staff who made my life so much easier - Sue Kellogg, Karen Grahl, Molly Martin and Alex Terzian. It was only with Alex's help that my parents and other family members could join my oral thesis defense remotely from India, it really meant a lot to me. I thank all HG faculty members who were always so nice to me. I thank Prof. Stephanie Bielas and Prof. Sundeep Kalantry for happily writing letters of recommendation for many applications. I thank all my Ann Arbor friends and fellow HG graduate students for all their support in so many ways.

I owe thanks to my teachers for their mentorship, especially Mrs. Shilpa Vishnoi and Mrs.Tajili Ansari during my early schooling; Prof. Aiyagari Ramesh and Prof. R. Swaminathan during my undergraduate studies and Prof. Philippe Renaud during my master's studies. They strongly motivated my interest in science and research and greatly supported me as I decided to pursue PhD studies and throughout the application process.  

My wonderful little family has been my driving force all these years. My sister Aakriti has been a lovely friend all along. My mother-in-law Mrs. Kanti Deshwali and sister-in-law Dr. Akanksha Deshwali have been nothing by warm and loving. My husband and best friend Dr. Anand Pratap Singh has unconditionally loved and supported me during the years and has been the best partner in all times good and bad. I thank him for being the amazingly calm person he is, listening to me and keeping me sane. My parents Mrs Beena and Mr. Pravish Varshney have always been unwaveringly supportive. Their love, encouragement and sacrifice has got me here today. I will strive to follow their examples of hard work and integrity in my career and life.